SVG: The Paper

### 2022-04-22 Analysis/index.Rmd
minor edits to rmd formatting

---
title: 'SVG: The Paper'
author: "Alasdair Clarke, Amelia Hunt &  Anna Hughes"
date: "13/04/2022"
output:
  tufte::tufte_html:
    toc: TRUE
    number_sections: true
    fig_height: 4 
---

@@

We will start by looking at the visual foraging data from Clarke et al (2022), QJEP. As the current version of our model only takes the $p_A$ bias into account for the initial target selection, we will ignore these data points for now.

```{r, echo = TRUE}
# read in data
d <- read_csv("data/clarke_2020_qjep.csv", show_col_types = FALSE) %>%
  mutate(condition = as_factor(condition),
         condition = fct_recode(condition, feature = "1", conjunction = "2"),
         targ_type = as_factor(targ_type))

# item weights pre-computed - see scripts/extract_item_weights_using_model.R
a <- readRDS("scratch/qjep_model_weights.rda") %>%
  filter(found != 1) # remove initial selections

@@

@@

From this figure we can see a few interesting things to note:

- large individual differences in how well our model can capture and predict behaviour
- feature search is more predictable than conjunction search
- our model appears to be overly conservative

@@

```{r, fig.cap = "Accuracy of our model varies with the strength of an individual's model parameters. We can see two clear outlier participants (marked with an X). "}
# computed in xxxx.R
fit <- readRDS("scratch/qjep_model_fit.rda") %>%
  full_join(a_acc, by = c("observer", "condition")) %>%
  mutate(is_outlier = if_else(observer %in% c(35, 45), TRUE, FALSE))

@@

ggplot(fit_plt, aes(value, accuracy, colour = condition)) + 
  geom_point(aes(shape = is_outlier)) +
  geom_smooth(data = filter(fit_plt, is_outlier == 0), method = "lm", formula = y ~ x, se = FALSE) + 
  scale_shape_manual(values = c(19, 4), guide = "none") + 
  facet_wrap(~parameter, scales = "free") +
  theme(legend.position = "bottom")
```
As the two outlier participants appear to be doing something different, not captured especially well by our model, we will remove them from the rest of this analysis.

@@

summary(lm(accuracy ~ condition * bP, 
           data = dm))

summary(lm(accuracy ~ bP + pS + bM + bP, 
           data = dm))

rm(dm)
```

@@

Let's score every trial by how well the model captures behaviour, and then plot some different examples

@@

```{r fig.height=3.5, fig.cap = "Histogram of how predictiable individual trials are, defined as how often the model assigned the largest weight to the target that was then selected."}
ggplot(a_trl, aes(prop_max)) + 
  geom_histogram(fill = "#34b6c6", breaks = seq(0, 1, 0.025)) +
  scale_x_continuous("trial predictability")
```

@@

```{r, fig.height=3.5, fig.cap = "(*left*) Histogram comparing the empirical path lengths to simulations from our fitted model. Difference indicates the human path length - the simulated path length. (*right*) Histogram comparing the median path lengths per participant."}
path_lengths <- readRDS("scratch/path_lengths.rda") %>%
  rename(observer = "obs",
         condition = "cond",
         human = "human_path_length",
         model = "model_path_length") %>% 
  mutate(difference  = human - model) %>%
  pivot_longer(c(human, model, difference), names_to = "type", values_to = "path_length")

>>>>

---
title: 'SVG: The Paper'
author: "Alasdair Clarke, Amelia Hunt &  Anna Hughes"
date: "13/04/2022"
output:
  tufte::tufte_html:
    number_sections: true
    fig_height: 4
    bibliography: literature.bib
    link-citations: yes  
---

@@

We will use the visual foraging data from @clarke2022 to assess the model's trial level performance. As the current version of our model only takes the $p_A$ bias into account for the initial target selection, we will ignore these data points for now.

```{r, echo = TRUE}
# read in data
d <- read_csv("data/clarke_2020_qjep.csv", show_col_types = FALSE) %>%
  mutate(condition = as_factor(condition),
         condition = fct_recode(condition, feature = "1", conjunction = "2"),
         targ_type = as_factor(targ_type))
```

We will use the model fit as detailed in @clarke2020compbio. For simplicity sake, we will characterise all of our estimated posterior probability densities with just the mean value. This gives us four parameters per participant: $b_S, b_A, b_P, b_D$.

Using these fitted values, we can now step through each target selection in the dataset and computer the probability weights our model assigns to each remaining target. For each target selection, we record:

- the item assigned the highest weight by our model.
- whether this item was then selected by the human participant.

```{r echo = TRUE}
# item weights pre-computed - see scripts/extract_item_weights_using_model.R
a <- readRDS("scratch/qjep_model_weights.rda") %>%
  filter(found != 1) # remove initial selections

@@

We can then plot this data to show how our ability to predict which target will be selected next varies throughout a trial. We can see a difference between *feature* and *conjunction* foraging, with *feature* foraging being more predictable.

@@

@@

```{r fig.height = 6, fig.cap = "Accuracy of our model varies with the strength of an individual's model parameters. We can see two clear outlier participants (marked with an X). "}
# computed in xxxx.R
fit <- readRDS("scratch/qjep_model_fit.rda") %>%
  full_join(a_acc, by = c("observer", "condition")) %>%
  mutate(is_outlier = if_else(observer %in% c(35, 45), TRUE, FALSE))

@@

ggplot(fit_plt, aes(value, accuracy, colour = condition)) + 
  geom_point(aes(shape = is_outlier)) +
  geom_smooth(data = filter(fit_plt, is_outlier == 0), method = "lm", formula = y ~ x, se = FALSE) + 
  scale_shape_manual(values = c(19, 4), guide = "none") + 
  facet_wrap(~parameter, scales = "free") +
  theme(legend.position = "bottom")
```

As the two outlier participants appear to be doing something different, not captured especially well by our model, we will remove them from the rest of this analysis.

@@

summary(lm(accuracy ~ condition * bP, 
           data = dm))
```

We can see that the difference in the proximity parameter between conditions appears to account for most of the variables in our model's accuracy from one participant to the next.

We can also see if the other parameters have an effect.

```{r echo=TRUE}
summary(lm(accuracy ~ pA + pS + bM + bP, 
           data = dm))
```

Both $p_A$ and $b_M$ also have an effect on our model's accuracy, but note there is only a small increases in $R^2$ from just using $b_P$.

```{r}
rm(dm)
```

@@

Let's score every trial by how well the model captures behaviour, and then plot some different examples.

@@

```{r fig.height=3.5, fig.cap = "Histogram of how predictiable individual trials are, defined as how often the model assigned the largest weight to the target that was then selected."}
ggplot(a_trl, aes(prop_max)) + 
  geom_histogram(fill = "#d04e00", breaks = seq(0, 1, 0.025), alpha = 0.5) +
  scale_x_continuous("trial predictability")
```

@@

```{r, fig.height=3.5, fig.cap = "(*left*) Histogram comparing the empirical path lengths to simulations from our fitted model. Difference indicates the human path length - the simulated path length. (*right*) Histogram comparing the median path lengths per participant.", message=FALSE}
path_lengths <- readRDS("scratch/path_lengths.rda") %>%
  rename(observer = "obs",
         condition = "cond",
         human = "human_path_length",
         model = "model_path_length") %>% 
  mutate(difference  = human - model) %>%
  pivot_longer(c(human, model, difference), names_to = "type", values_to = "path_length")
